---
title: "task 1"
author: "Mari Herbst"
date: "1/28/2022"
output: html_document
code_folding: hide
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# attach packages
library(tidyverse)
library(here)
library(patchwork)
library(broom)
library(kableExtra)
```

# Introduction

# Data

## Citation

Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5

## Visualizations

```{r}
palmetto <- read_csv(here("data", "palmetto.csv")) %>% # read in data using here() to create a system agnostic file path
  mutate(species = ifelse(species == 1, "repens", "etonia")) %>% # change species number to species name for clarity
  select(species, height, length, width, green_lvs) %>%
  drop_na()

palmetto$species <- as.factor(palmetto$species) # change species from character to factor
```

Factor ranking:

species 1: Sabal etonia
species 2: Serenoa repens

Some visualizations to explore relationships between variables.

```{r}
length_height <- ggplot(data = palmetto, aes(x = length, y = height)) +
  geom_point(aes(color = species), show.legend = FALSE) + # color by species, remove legend
  facet_wrap(~species) + # 2 plots by species
  scale_color_manual(values = c("etonia" = "darkolivegreen", "repens" = "darkolivegreen2")) + # set color for each species
  labs(x = "length (cm)", # add axis labels and title
       y = "height (cm)",
       title = "(a) Comparison of Serenoa repens and Sabal etonia palmetto length and height") +
  theme(plot.title = element_text(hjust = 0.5, # adjust title position and size
                                  size = 10.5))
```

```{r}
leaves_width <- ggplot(data = palmetto, aes(x = green_lvs, y = width)) +
  geom_point(aes(color = species)) + # color by species
  scale_color_manual(values = c("etonia" = "darkolivegreen", "repens" = "darkolivegreen2")) + # set color for each species
  labs(x = "# green leaves", # label axes and title
       y = "width (cm)",
       title = "(b) Comparison of Serenoa repens and Sabal etonia palmetto width and leafiness") +
  theme(plot.title = element_text(hjust = 0.3, # adjust title position and size
                                  size = 10.5))
```

```{r}
species_leaves <- ggplot(data = palmetto, aes(x = species, y = green_lvs)) +
  geom_jitter(aes(color = species), show.legend = FALSE) + # color by species, remove legend
  scale_color_manual(values = c("etonia" = "darkolivegreen", "repens" = "darkolivegreen2")) + # set color for each species
  labs(x = "species", # add axes labels and title
       y = "# green leaves",
       title = "(c) Comparison of Serenoa repens and Sabal etonia palmetto leafiness") +
  theme(plot.title = element_text(hjust = 0.5, # adjust title position and size
                                  size = 10.5))
```

```{r}
# combine visualizations using patchwork
length_height + leaves_width + species_leaves + plot_layout(ncol = 1, heights = c(8,8,8)) # set number of columns and size of graphs
```


**Figure 1**: *(a)* An exploration of canopy lengths and plant heights of *Serenoa repens* (light green) and *Sabal etonia* (dark green), two species of palmetto. *(b)* A comparison of canopy width versus leafiness of *Serenoa repens* and *Sabal etonia* palmettos. *(c)* Shows the relative leafiness of the two palmetto species.

From comparing these variables and analyzing the visualizations, it appears as though leafiness may be a distinct characteristic between the two species. *Serenoa repens* also seems to trend somewhat taller, while *Sabal etonia* trends somewhat longer.

# Binary logistic regression models

```{r}
# create the formulas
f1 <- species ~ height + length + width + green_lvs
f2 <- species ~ height + width + green_lvs

# create the models
palm_blr1 <- glm(formula = f1,
                 data = palmetto,
                 family = "binomial")

palm_blr2 <- glm(formula = f2,
                 data = palmetto,
                 family = "binomial")

blr1_tidy <- broom::tidy(palm_blr1)
```

## AICc

```{r}
# create a table to compare AICcs of the models
aic <- AICcmodavg::aictab(list(palm_blr1, palm_blr2))

# round table values to a reasonable number of digits
aic$AICc[1] <- round(aic$AICc[1], 2)
aic$AICc[2] <- round(aic$AICc[2], 2)
aic$Delta_AICc[1] <- round(aic$Delta_AICc[1], 2)
aic$Delta_AICc[2] <- round(aic$Delta_AICc[2], 2)
aic$ModelLik[1] <- round(aic$ModelLik[1], 2)
aic$ModelLik[2] <- round(aic$ModelLik[2], 2)
aic$AICcWt[1] <- round(aic$AICcWt[1], 2)
aic$AICcWt[2] <- round(aic$AICcWt[2], 2)
aic$LL[1] <- round(aic$LL[1], 2)
aic$LL[2] <- round(aic$LL[2], 2)
aic$Cum.Wt[1] <- round(aic$Cum.Wt[1], 2)
aic$Cum.Wt[2] <- round(aic$Cum.Wt[2], 2)

# finalize table
aic %>%
  kable(col.names = c("model", "K", "AICc", "change in AICc", "model likelihood", "Akaike weight", "log likelihood", "cumulative Akaike weight")) %>% # finalize column names
  kable_paper() # apply theme
```

## Cross validate

10-fold cross-validation to compare model 1 to model 2

```{r}
# use caret package to automate cross validation
set.seed(615) # set seed so same random numbers will be used for reproducibility

train_ctrl <- trainControl(method = "repeatedcv",
                           number = 10, # number of folds
                           repeats = 10) # number of iterations

# train for model 1
model1 <- train(f1, data = palmetto,
                method = "glm",
                family = "binomial",
                trControl = tr_ctrl)

# train for model 2
model2 <- train(f2, data = palmetto,
                method = "glm",
                family = "binomial",
                trControl = tr_ctrl)


